---
title: "Collagen and related proteins list for Nordic"
output: html_notebook
---

Load required packages.
```{r}
library(devtools)
mypackages <- c("ggplot2", "tidyverse")
lapply(mypackages, require, character.only = TRUE)
library(SomaDataIO)
```

Read in data and check format.
```{r}
file <- "~/Desktop/DRG/Proteomics/10_28_22_Merged_With_Legend_SS-2216784_v4.1_EDTAPlasma.hybNorm.medNormInt.plateScale.calibration.anmlQC.qcCheck.anmlSMP.adat"
adat <- read_adat(file)
is.soma_adat(adat)
adat
```

The first part of this notebook is assessing other collagen entities.

Subset df to only desired columns and rename to Entrez gene ID for ease of discrimination.
```{r}
collagen_list <- c('seq.11140.56',
'seq.13484.69',
'seq.15466.30',
'seq.6631.17',
'seq.22047.46',
'seq.11150.3',
'seq.16828.8',
'seq.4807.13',
'seq.15653.9',
'seq.6570.1',
'seq.8974.172',
'seq.8804.39',
'seq.4543.65',
'seq.7006.4',
'seq.10702.1',
'seq.16753.46',
'seq.11278.4',
'seq.20175.17',
'seq.11196.31',
'seq.10511.10',
'seq.11155.16',
'seq.15467.10',
'seq.6236.51',
'seq.15569.15',
'seq.18880.81')

collagen_names <- c('COL1A1.1',
'COL1A1.2',
'COL9A1.1',
'COL9A1.2',
'COL5A1',
'COL6A1.1',
'COL6A1.2',
'COL8A1',
'COL10A1',
'COL13A1',
'COL15A1',
'COL20A1',
'COL23A1',
'COL25A1',
'COL28A1',
'COL6A2',
'COL11A2',
'COL9A3',
'COL6A3.1',
'COL6A3.2',
'COL6A5',
'CTHRC1.1',
'CTHRC1.2',
'COL2A1',
'COL3A1')

sub_adat <- adat %>% select('Cohort', all_of(collagen_list))
sub_adat <- sub_adat %>% rename_at(vars(collagen_list), ~ collagen_names)
sub_adat
```
Perform t-tests between Control and Dupuytren conditions for each protein.
```{r}
Cohort <- sub_adat$Cohort
res <- sub_adat %>% 
  select_if(is.numeric) %>%
  map_df(~ broom::tidy(t.test(. ~ Cohort)$p.value), .id = 'Protein')
res %>% rename('p-value' = 'x')
```

Loop over df and visualize data.
```{r}
for(i in collagen_names) {
  print(ggplot(sub_adat, aes(x=Cohort)) + geom_boxplot(aes_string(y= i)))
}
```

The final part of this notebook is assessing other collagen related and modifying proteins (among some other collagen proteins as well).

Subset df to only desired columns and rename to Entrez gene ID for ease of discrimination.
```{r}
protein_list <- c("seq.10479.18", "seq.10511.10","seq.10612.18", "seq.10800.15","seq.11150.3","seq.11196.31","seq.11237.49","seq.11348.132","seq.11645.9","seq.13535.2","seq.13950.9"  ,"seq.15466.30","seq.15569.15","seq.16753.46","seq.16828.8","seq.18875.125","seq.18880.81","seq.20175.17","seq.2201.17","seq.22047.46","seq.2579.17","seq.2788.55"  ,"seq.2789.26","seq.2954.56","seq.3348.49","seq.4160.49","seq.4496.60","seq.4543.65","seq.4924.32","seq.4925.54","seq.5002.76","seq.5638.23","seq.6273.58",
"seq.6383.90","seq.6570.1","seq.6631.17","seq.8475.15","seq.8479.4" ,"seq.8845.2" ,"seq.9172.69")

protein_names <- c("MMP10_1","COL6A3_1","PLOD3","SERPINH1","COL6A1_1","COL6A3_2","PCOLCE","P4HA2","P4HA1","CERT1_1","CERT1_2","COL9A1_1","COL2A1_1","COL6A2","COL6A1_2","COL2A1_2","COL3A1","COL9A3","COL18A1","COL5A1","MMP9","MMP3","MMP7_1","MMP8_1","BMP1","MMP2","MMP12","COL23A1","MMP1","MMP13",
"MMP14","COLGALT1","P3H1","TLL1","COL13A1","COL9A1_2","MMP7_2","MMP10_2","ADAMTS3","MMP8_2")

sub_adat2 <- adat %>% select('Cohort', all_of(protein_list))
sub_adat2 <- sub_adat2 %>% rename_at(vars(protein_list), ~ protein_names)
sub_adat2
```

Perform t-tests between Control and Dupuytren conditions for each protein.
```{r}
Cohort <- sub_adat2$Cohort
res <- sub_adat2 %>% 
  select_if(is.numeric) %>%
  map_df(~ broom::tidy(t.test(. ~ Cohort)$p.value), .id = 'Protein')
res %>% rename('p-value' = 'x')
```

Loop over df and visualize data.
```{r}
for(i in protein_names) {
  print(ggplot(sub_adat2, aes(x=Cohort)) + geom_boxplot(aes_string(y= i)))
}
```
